%% =============================================================================
%% sm_c3_bvm.tex — SM-C, Subsection C.3: Bernstein–von Mises Approximation
%% Body content only (\subsection level); parent structure in sm_c.tex
%% Primary source: dev/log/06_research-notes-survey-design-pseudo-posterior.qmd
%%   lines 784--922 (Theorem 3.2, Proposition 3.8, Corollary 3.1, Remarks)
%% Correction applied: T1-1 (BvM reframe — diagnostic, not formal violation)
%% =============================================================================

\subsection{Bernstein--von Mises approximation}
\label{smc:bvm}

The sandwich correction in~\cref{sec:survey} rests on a
Bernstein--von Mises (BvM) theorem for pseudo-posteriors: as
$N \to \infty$, the pseudo-posterior converges to a normal distribution
centered at the pseudo-MLE with variance given by the inverse observed
Hessian, $\mathbf{H}_{\mathrm{obs}}^{-1}$.  We state the precise result, verify its regularity
conditions for the HBB model, and provide a finite-sample
diagnostic for the NSECE weight configuration that motivates the
sandwich correction as a robustness measure.

%% --- BvM Theorem ---

\begin{theorem}[Design-consistent BvM for the pseudo-posterior]
\label{smc:thm-bvm}
  Suppose the following regularity conditions hold\textup{:}
  \begin{enumerate}[label=\textup{(C\arabic*)},nosep]
    \item \label{smc:C1}%
      \textup{Smoothness.}
      The log-density $\log f_{\HBB}(y \mid \btheta)$ is twice
      continuously differentiable in $\btheta$ on the interior of the
      parameter space $\Theta$.
    \item \label{smc:C2}%
      \textup{Interior parameter.}
      The true superpopulation parameter $\btheta_0$ lies in the
      interior of $\Theta$.
    \item \label{smc:C3}%
      \textup{Weight regularity.}
      The ratio of the largest to smallest normalized weight satisfies
      $\max_i \tilde{w}_i / \min_i \tilde{w}_i = O(N^{\delta})$ for
      some $\delta < 1/2$ as $N \to \infty$.
    \item \label{smc:C4}%
      \textup{Information regularity.}
      The weighted Hessian $\mathbf{H}_{\mathrm{obs}}(\btheta_0)$ is
      positive definite, with eigenvalues bounded away from zero and
      infinity.
    \item \label{smc:C5}%
      \textup{Prior regularity.}
      The prior density $\pi(\btheta)$ is continuous and strictly
      positive in a neighborhood of $\btheta_0$.
  \end{enumerate}
  Then, as $N \to \infty$\textup{,}
  \begin{equation}\label{smc:eq-bvm}
    \sup_{B \in \cl{B}}
    \bigl|
      \pi^{(w)}(\btheta \in B \mid \by)
      - \Phi_{\hat{\btheta}^{(w)},\,\mathbf{H}_{\mathrm{obs}}^{-1}}(B)
    \bigr|
    \xrightarrow{P_{\btheta_0}} 0,
  \end{equation}
  where $\cl{B}$ is the class of Borel-measurable convex sets,
  $\hat{\btheta}^{(w)}$ is the pseudo-MLE,
  $\Phi_{\bmu,\bSigma}$ denotes the CDF of
  $\Norm(\bmu, \bSigma)$, and $\mathbf{H}_{\mathrm{obs}}$ is the
  observed Hessian of the pseudo-log-likelihood evaluated at
  $\hat{\btheta}^{(w)}$.
\end{theorem}

The result is a pseudo-posterior analogue of the classical BvM theorem
\citep{KleijnVanDerVaart2012}, extended to the survey-weighted setting by
\citet{SavitskyToth2016} and~\citet{WilliamsSavitsky2021}.  The key
difference from the standard BvM is condition~\ref{smc:C3}, which
constrains how rapidly the weight ratio may grow with sample size.
Importantly, the \emph{pseudo-posterior variance}
$\mathbf{H}_{\mathrm{obs}}^{-1}$ differs from the
\emph{design-consistent sampling variance} of $\hat{\btheta}^{(w)}$,
\[
  \hat{\btheta}^{(w)}
  \;\dot{\sim}\;
  \Norm\!\bigl(\btheta_0,\;
  \mathbf{V}_{\mathrm{sand}}\bigr),
  \qquad
  \mathbf{V}_{\mathrm{sand}}
  = \mathbf{H}_{\mathrm{obs}}^{-1}\,
    \mathbf{J}_{\mathrm{cluster}}\,
    \mathbf{H}_{\mathrm{obs}}^{-1},
\]
as defined in~\eqref{eq:V-sand}. Under informative sampling or model
misspecification,
$\mathbf{J}_{\mathrm{cluster}} \neq \mathbf{H}_{\mathrm{obs}}$,
so naive credible intervals based on $\mathbf{H}_{\mathrm{obs}}^{-1}$
are miscalibrated as a measure of frequentist uncertainty.  This
mismatch is precisely what motivates the sandwich correction developed
in~\cref{smc:sandwich}: \cref{smc:cor-coverage} below quantifies
the resulting coverage distortion, and
\cref{thm:cholesky} provides the operational calibration.

%% --- Verification of conditions ---

\paragraph{Verification.}
We verify each condition for the HBB model and the NSECE data.

\medskip\noindent\textit{\ref{smc:C1}: Smoothness.}\enspace
The HBB log-density involves $\log\Gamma$ functions (through the
beta-binomial PMF), the logistic link
$\sigma(\eta) = 1/(1+e^{-\eta})$, and the zero probability $p_0$
(a finite sum of $\log\Gamma$ terms).  Each component is $C^{\infty}$
on the interior of the parameter space, so the composite
$\log f_{\HBB}$ is smooth in $\btheta$.  \checkmark

\medskip\noindent\textit{\ref{smc:C2}: Interior parameter.}\enspace
The parameter space is
$\Theta = \R^P \times \R^P \times \R^{2QR} \times
\mathbb{S}_{++}^{2Q} \times \R_{>0}$,
and any finite parameter vector with $\kappa > 0$ and
$\bSigma_\varepsilon \succ 0$ lies in the interior.  \checkmark

\medskip\noindent\textit{\ref{smc:C3}: Weight regularity.}\enspace
This condition requires a dedicated analysis; see
\cref{smc:prop-weight-diagnostic} below.

\medskip\noindent\textit{\ref{smc:C4}: Information regularity.}\enspace
By~\cref{smb:prop-block-diag}, the Fisher information is
block-diagonal between the extensive and intensive margins.  For the
extensive block, positive definiteness follows from $\operatorname{rank}(\mathbf{X}) = P$
and $q_i \in (0,1)$ for all $i$ (standard logistic regression theory).
For the intensive block, separation of $\mu$ from $\kappa$ requires
variation in the denominators $n_i$
(\cref{smb:prop-mu-kappa-id}), which is satisfied empirically
($n_i$ ranges from~1 to~378 among positive responders).  \checkmark

\medskip\noindent\textit{\ref{smc:C5}: Prior regularity.}\enspace
The priors specified in~\cref{sec:priors} are all continuous and
strictly positive in the interior:
$\Norm(\mathbf{0},\, 2^2\,\mathbf{I}_P)$ on regression coefficients~\citep{GelmanEtAl2008},
$\HalfNormal(0,1)$ on scale parameters,
$\LKJ(2)$ on the correlation matrix~\citep{LewandowskiEtAl2009}, and
$\Norm(2,\, 1.5^2)$ on $\log\kappa$.  \checkmark

%% --- Finite-sample weight diagnostic ---

\begin{proposition}[Finite-sample weight diagnostic]
\label{smc:prop-weight-diagnostic}
  For the NSECE 2019 data with $N = 6{,}785$ and
  $w_{\max}/w_{\min} = 462$\textup{,} the weight ratio exceeds
  $N^{1/2} \approx 82.4$\textup{,} indicating that the sufficient
  condition \ref{smc:C3} is not comfortably satisfied for this design.
  The computation
  \begin{equation}\label{smc:eq-delta}
    \delta
    = \frac{\log 462}{\log 6{,}785}
    = \frac{6.136}{8.822}
    \approx 0.695
    > 0.5
  \end{equation}
  is a finite-sample diagnostic, not a valid asymptotic big-$O$ test
  of condition \ref{smc:C3}.  This motivates the sandwich variance
  correction as a robustness measure.
\end{proposition}

\begin{remark}[Interpretation of the diagnostic]
\label{smc:rem-weight-interpretation}
  Four observations clarify the scope of
  \cref{smc:prop-weight-diagnostic}.
  \begin{enumerate}[label=\textup{(\alph*)},nosep]
    \item Condition~\ref{smc:C3} is \emph{sufficient}, not necessary.
      The BvM approximation may remain adequate even when this bound is
      not met, provided the high-weight observations do not dominate
      the pseudo-likelihood.
    \item The fraction of observations with extreme weights is small:
      the top $1\%$ (weights $w_i > 100$) comprises fewer than 70
      observations out of $6{,}785$.  The diagnostic is driven by a
      small number of outlying weights, not a systematic pattern.
    \item Weight trimming can restore~\ref{smc:C3}.  Capping at the
      99th percentile reduces $w_{\max}$ to approximately~100, giving
      $\delta \approx 0.522$; capping at the 95th percentile
      ($w_{\max} \approx 50$) yields $\delta \approx 0.443 < 0.5$,
      restoring \ref{smc:C3}.  Weight trimming is examined as a
      sensitivity analysis in~\cref{app:extended-results}.
    \item In practice, the sandwich correction
      \eqref{eq:V-sand} provides the appropriate variance adjustment
      regardless of whether~\ref{smc:C3} holds exactly: it
      automatically accounts for the inflated variance due to extreme
      weights.
  \end{enumerate}
\end{remark}

%% --- Coverage calibration ---

\begin{corollary}[Coverage miscalibration of naive credible intervals]
\label{smc:cor-coverage}
  Under the conditions of~\cref{smc:thm-bvm}, define\textup{:}
  \begin{itemize}[nosep]
    \item $\mathrm{CI}_{\mathrm{naive}}(\alpha)$\textup{:} the
      $\alpha$-level credible interval from the pseudo-posterior
      marginal, with variance
      $[\mathbf{H}_{\mathrm{obs}}]^{-1}$\textup{;}
    \item $\mathrm{CI}_{\mathrm{sand}}(\alpha)
      = \hat{\theta}^{(w)} \pm z_{\alpha/2}\sqrt{V_{\mathrm{sand}}}$\textup{:}
      the sandwich-corrected interval.
  \end{itemize}
  Then\textup{:}
  \begin{enumerate}[label=\textup{(\alph*)},nosep]
    \item The naive interval has asymptotic coverage
      $\Phi(z_{\alpha/2}\, r) - \Phi(-z_{\alpha/2}\, r)$, where
      $r = \sqrt{\lambda_{\mathbf{H}} / \lambda_{\mathrm{sand}}}$ is
      the ratio of the naive to the sandwich standard deviation and
      $\Phi$ denotes the standard normal CDF.
    \item When the sampling is non-informative and the model is
      correctly specified, $\mathbf{J}_{\mathrm{cluster}} \approx
      \mathbf{H}_{\mathrm{obs}}$, so $r \approx 1$ and the naive
      interval has approximately correct coverage.
    \item Under informative sampling or model misspecification,
      $r \neq 1$.  If $r < 1$ \textup{(}naive variance too
      small\textup{)}, the interval is anti-conservative
      \textup{(}undercoverage\textup{)}---the typical case when the
      pseudo-posterior overconcentrates.  If $r > 1$, the interval is
      conservative.
  \end{enumerate}
\end{corollary}

\begin{proof}
  The BvM theorem (\cref{smc:thm-bvm}) implies that the
  pseudo-posterior converges to
  $\Norm(\hat{\btheta}^{(w)},\, \mathbf{H}_{\mathrm{obs}}^{-1})$,
  while the correct sampling distribution of $\hat{\btheta}^{(w)}$ is
  approximately $\Norm(\btheta_0,\, \mathbf{V}_{\mathrm{sand}})$.
  The ratio of marginal standard deviations determines the coverage
  distortion.
\end{proof}

\begin{remark}[Calibration via the Cholesky transformation]
\label{smc:rem-cholesky-calibration}
  \cref{thm:cholesky} provides the operational calibration
  strategy: the affine transformation
  \eqref{eq:cholesky-transform} rescales each MCMC draw so that
  the adjusted posterior has covariance
  $\mathbf{V}_{\mathrm{sand}}$ while preserving the posterior mean.
  This replaces the naive variance $\mathbf{H}_{\mathrm{obs}}^{-1}$
  with the design-consistent sandwich variance, correcting the
  coverage distortion quantified in~\cref{smc:cor-coverage}.  The
  block-wise implementation---applying the transformation separately
  to fixed effects, random effects, and hyperparameters---is discussed
  in~\cref{sec:survey}; the full proof appears in
  \cref{smc:cholesky} below.
\end{remark}
