%% =============================================================================
%% sm_f1_dgp.tex — SM-F, Subsection F.1: Data-Generating Process
%% Body content only (\subsection level); parent structure in sm_f.tex
%% =============================================================================

\subsection{Data-generating process and calibration}
\label{smf:dgp}

This section provides the complete data-generating process (DGP)
for the simulation study in~\cref{sec:simulation}.
The DGP is designed to match the NSECE 2019 data structure while
allowing parametric control over survey informativeness.

\paragraph{Population generation.}
A finite super-population of $M = 50{,}000$ providers is distributed
across $S = 51$ states, with state sizes proportional to the
observed NSECE state counts.  For each provider~$i$ in state~$s$:
\begin{enumerate}
  \item Draw the covariate vector
    $\bx_i = (1,\, x_{\mathrm{pov}},\, x_{\mathrm{urb}},\,
    x_{\mathrm{blk}},\, x_{\mathrm{his}})^\t$ by bootstrap
    resampling from the empirical NSECE marginals within state~$s$.
  \item Draw the trial size $n_i$ from the empirical distribution
    of total 0--5 enrollment in the NSECE data.
\end{enumerate}

\paragraph{True parameter values.}
The true parameters are calibrated to the M1 (random intercept)
unweighted NSECE fit (\cref{sme:model-hierarchy}).
\Cref{smf:tab-truth} reports all parameter values used in the
DGP.  These values were chosen to produce a realistic structural
zero rate (approximately 36\%) and overdispersion level
consistent with the empirical data.

\begin{table}[htbp]
  \centering
  \small
  \begin{tabular}{@{}l r l@{}}
    \toprule
    Parameter & Value & Description \\
    \midrule
    $\alpha_{\text{int}}$     & $+0.696$ & Extensive intercept \\
    $\alpha_{\text{pov}}$     & $-0.119$ & Extensive poverty \\
    $\alpha_{\text{urb}}$     & $+0.253$ & Extensive urban \\
    $\alpha_{\text{blk}}$     & $-0.070$ & Extensive Black \\
    $\alpha_{\text{his}}$     & $-0.139$ & Extensive Hispanic \\
    \addlinespace
    $\beta_{\text{int}}$      & $-0.032$ & Intensive intercept \\
    $\beta_{\text{pov}}$      & $+0.057$ & Intensive poverty \\
    $\beta_{\text{urb}}$      & $-0.018$ & Intensive urban \\
    $\beta_{\text{blk}}$      & $+0.080$ & Intensive Black \\
    $\beta_{\text{his}}$      & $+0.040$ & Intensive Hispanic \\
    \addlinespace
    $\log\kappa_0$            & $+1.655$ & Log concentration \\
    $\kappa_0$                & $5.232$ & Concentration \\
    \addlinespace
    $\tau_{\text{ext}}$       & $0.577$ & State SD (extensive) \\
    $\tau_{\text{int}}$       & $0.208$ & State SD (intensive) \\
    $\rho_{\text{cross}}$     & $0.285$ & Cross-margin correlation \\
    \bottomrule
  \end{tabular}
  \caption{True parameter values for the simulation DGP, calibrated to
    the M1 unweighted fit.  Fixed effects are on the logit scale;
    $\kappa_0 = \exp(\log\kappa_0)$.}
  \label{smf:tab-truth}
\end{table}

\paragraph{Outcome generation.}
For each replication~$r = 1,\ldots,R$:
\begin{enumerate}
  \item Draw state random effects
    $\bdelta_s = (\delta_{1s},\, \delta_{2s})^\t \sim
    \Norm_2(\mathbf{0}, \bSigma_\delta)$, where
    \[
      \bSigma_\delta = \diag(\boldsymbol{\tau})\,
      \mathbf{R}_\delta\,\diag(\boldsymbol{\tau}),
      \quad
      \boldsymbol{\tau} = (\tau_{\text{ext}},\,\tau_{\text{int}})^\t,
      \quad
      [\mathbf{R}_\delta]_{12} = \rho_{\text{cross}}.
    \]
  \item Compute linear predictors:
    $\eta_i^{\text{ext}} = \bx_i^\t\balpha_0 + \delta_{1,s[i]}$,
    $\eta_i^{\text{int}} = \bx_i^\t\bbeta_0 + \delta_{2,s[i]}$.
  \item Generate participation:
    $z_i \sim \Bern(q_i)$ with
    $q_i = \expit(\eta_i^{\text{ext}})$.
  \item For providers with $z_i = 1$, generate positive counts via
    \textbf{rejection sampling} from the unconditional beta-binomial
    \citep{Prentice1986}:
    \begin{quote}
      Draw $Y_i \sim \BetaBin(n_i,\, \mu_i,\, \kappa_0)$ where
      $\mu_i = \expit(\eta_i^{\text{int}})$;
      \emph{repeat until $Y_i > 0$}; set $y_i = Y_i$.
    \end{quote}
    This ensures $y_i > 0$ for all providers with $z_i = 1$,
    consistent with the zero-truncated beta-binomial component of
    the hurdle model.  The acceptance probability is $1 - p_0(n_i,
    \mu_i, \kappa_0)$, which exceeds~$0.80$ for the vast majority
    of observations in our calibration.
  \item For providers with $z_i = 0$, set $y_i = 0$.
\end{enumerate}

\begin{remark}[T3-3: Zero-truncated beta-binomial DGP]
  \label{smf:rem-ztbb}
  Drawing from the unconditional $\BetaBin$ and rejecting zeros
  produces exact draws from the zero-truncated beta-binomial
  $\ZTBB(n, \mu, \kappa)$.  The alternative---dividing the
  unconditional PMF by $(1-p_0)$---would require a custom
  sampling routine, whereas rejection sampling reuses the
  standard $\BetaBin$ sampler and is efficient whenever $p_0$ is
  not too close to~$1$.  The same approach is used in the
  Stan \texttt{generated quantities} block for posterior
  predictive replication.
\end{remark}

\paragraph{Sampling design.}
From each finite population, a stratified cluster sample of
approximately $N \approx 7{,}000$ providers is drawn using
size-biased inclusion probabilities
$\log\pi_i = c_0 + \rho_{\mathrm{inc}} \cdot y_i^*$
(\cref{eq:sim-inclusion}), where $y_i^*$ is the standardized
latent outcome and $c_0$ is calibrated to the target sample size
\citep[cf.][]{SavitskyToth2016}.
The three scenarios (S0, S3, S4) are defined in
\cref{tab:sim-scenarios}; see~\cref{sec:simulation} for the
full discussion.  Survey weights are computed as
$w_i = 1/\pi_i$ and normalized to sum to~$N$ for the
pseudo-posterior.
